The quench action approach for integrable models: A Monte Carlo implementation

### 2015-09-24 draft/fs_qa_v2.tex
rev-24/09/2015

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel state overlap}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The string hypothesis: Reduced N\'eel overlap formulas}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overlap with the Majumdar-Ghosh state}

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
The overlap between a generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state can be obtained from the N\'eel state overlap in Eq.~\eref{neel-ov} 
as~\cite{pozsgay-2014} 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\pm\lambda_j\}_{j=1}^m\rangle=\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)
\langle N|\{\pm\lambda_j\}_{j=1}^m\rangle
\end{equation}
%
The mutliplicative factor in~\eref{mg-ov}, using the string hypothesis is 
rewritten as 
%
\begin{eqnarray}
\label{mg-k}
\fl\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)=\\\nonumber\qquad
2^m\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}
\lambda_{j;\ell}^{1+(-1)^j}(\lambda^2_{j;\ell}+j^2)\prod\limits_{k=0}^{\lfloor 
j/2\rfloor}\Big[\lambda_{j;\ell}^2+\Big(2k+\frac{1-(-1)^j}{2}\Big)^2\Big]^{-2}
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The N\'eel overlap in the thermodynamic limit}

@@

The functional integral in~\eref{qa-d-ensemble} can be evaluated in the limit 
$L\to\infty$ using the saddle point approximation. One has to minimize the 
functional ${\mathcal F}(\pmb{\rho})$ defined as 
%
\begin{equation}
L{\mathcal F}(\pmb{\rho})\equiv 2\Re\log\langle\pmb{\rho}|\Psi_0
\rangle+S_{YY}(\pmb{\rho}(\lambda))  
\end{equation}
% 
with respect to $\pmb{\rho}$, under the constraint that the thermodynamic BGT 
equations~\eref{bgt-th} hold. Remarkably, for the quench with initial 
state the N\'eel state $|\Psi_0\rangle=|N\rangle$ the resulting saddle point root 
distributions $\rho_n^*(\lambda)$ can be obtained analytically~\cite{brockmann-2014}. 
The first few are given as 
%
\begin{eqnarray}
\label{rho1-sp}
\fl\quad\rho^*_1(\lambda)=\frac{8(4+\lambda^2)}{\pi(19+3\lambda^2)(1+6\lambda^2+
\lambda^4)}\\\label{rho2-sp}
\fl\quad\rho_2^*(\lambda)=\frac{8\lambda^2(9+\lambda^2)(4+3\lambda^2)}{\pi(2+\lambda^2)
(16+14\lambda^2+\lambda^4)(256+132\lambda^2+9\lambda^4)}\\\label{rho3-sp}
\fl\quad\rho_3^*=\frac{8(1+\lambda^2)^2(5+\lambda^2)(16+\lambda^2)(21+\lambda^2)}
{\pi(19+3\lambda^2)(9+624\lambda^2+262\lambda^4+32\lambda^6+\lambda^8)
(509+5\lambda^2(26+\lambda^2))}.
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{N\'eel overlaps: Numerical results and overview}
\label{sec:5}

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel_overlaps}
\end{center}
\caption{ N\'eel state overlaps in the Heisenberg spin chain. (a) 
 The N\'eel state overlaps with the eigenstates of the Heisenberg 
 spin chain: Squared overlaps $|\langle\lambda|N\rangle|^2$ plotted 
 as a function of the eigenstates energy density $E/L$. Here $|\lambda
 \rangle$ denotes the generic eigenstate of the Heisenberg spin chain. 
 The data are for chains with length $26\le L\le 38$. The data are 
 obtained from a full scanning of the chains Hilbert space. 
 Eigenstates containing strings with zero-momentum components 
 are escluded. The overlaps decay exponentially with the chain size. 
 (b) Histograms of $-2\log|\langle\lambda|N\rangle|/L$. Different 
 histograms are for different chain sizes. The histogram bin width 
 is $\sim 2/L$. The $y$-axis is rescaled by a factor $10^5$ for 
 convenience. Note the peaking of the histograms around 
 $-2\log|\langle\lambda|N\rangle|\sim 0.5$ in the thermodynamic 
 limit {\bf really??}. The dotted vertical line is the expected 
 Quench-Action result in the thermodynamic limit.  
}
\label{fig0:neel-ov}
\end{figure}
%##################################################################

In this section we present the generic features of the overlaps between the Bethe 
states of the Heisenberg spin chain and the N\'eel state. We focus on finite chains 
of length up to $L\lesssim 40$. We consider the full Hilbert space of the chain, 
i.e., all the eigenstates. These are obtained from the Bethe ansatz approach. 
We exclude all the eigenstates corrsponding to BGT roots containing zero-momentum 
strings. The number of remaining eigenstates, that we denote as $\widetilde Z_{Neel}$ 
is given as a function of the chain size in terms of simple combinatorial formualas 
that we provide.

Interestingly, we demonstrate that for each finite chain the fraction of the 
eigenstates with no zero-momentum strings is vanishing in the thermodynamic limit.

We restrict ourselves to eigenstates corresponding to BGT roots with no 
zero-momentum strings. As expected the N\'eel overlaps decay exponentially with 
the chain size in the thermodynamic limit.

Moreover, all the overlaps exhibit a vanishing behavior as a function of the eigenstate 
energy density, signalling that most of the N\'eel overlap is concentrated at low 
energies.

We observe that the vast majority of eigenstates have negligible N\'eel overlap, 
as expected.

We show that only a small fraction of the eigenstates are relevant in the 
thermodynamic limit.

These are the Bethe states that are relevant in the quench action approach.

Finally, we focus on several overlap sum rules. As a consequence of the fact that we 
neglect the zero-momentum strings, we observe a striking violation of the sum rules 
for any finite chain. Specifically, all the sum rules exhibit vanishing behavior 
as $1/\sqrt{L}$ upon increasing the chain size, reflecting the vanishing of the 
fraction of the eigenstates with no zero-momentum strings.

Similarly, for the Majumdar-Ghosh state the sum rules vanish as $1/L$, again, 
reflecting the vanishing behaivor of the fraction of eigenstates with no 
zero-momentum strings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel overlaps from the  full scanning of the Hilbert space}
\label{sec:5.1}

Here we present numerical results for the overlaps between the eigenstates of the 
Heisenberg spin chain and the N\'eel state. The data correspond to the full set 
of eigenstates with no zero-momentum strings and chain lengths $L\le 38$. Their 
number $\widetilde Z_{Neel}$ is given as (see~\ref{app-2} for the proof) 
%
\begin{equation}
\label{ztilde}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big).
\end{equation}
%
For $L=38$ from~\eref{ztilde} one has $\widetilde Z_{Neel}\sim 10^5$ eigenstates. 
Note that the total number of partiy-invariant eigenstates $Z_{Neel}$, which 
in principle give non-zero N\'eel overlap, is given as 
%
\begin{equation}
\label{zneel1}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1.
\end{equation}
%
The proof of~\eref{zneel1} is reported in~\ref{app-1}.

We should stress that $Z_{Neel}$ provides only an upper bound for the number of 
eigenstates with non-zero N\'eel overlap (cf. for instance Table~\ref{table:neel}). 
This is due to the fact that parity-invariant eigenstates with a single even-length 
string with zero-momentum, which are counted in $Z_{Neel}$, have vanishing N\'eel 
overlap. This is due to the fact that in~\eref{neel-ov} the ratio of determinants 
is finite, whereas the prefactor (cf.~\eref{neel-k}) vanishes if a single even-length 
string with zero momentum is present.

{\bf Clarify what is the number/fraction of the even-length zero-momentum strings}

The results are obtained by generating all the parity-invariant BGT quantum numbers 
and using them to solve the BGT equations~\eref{bgt-eq}, obtaining the rapidities of 
the parity-invariant eigenstates of the $XXX$ chain. The overlaps with the N\'eel 
states are calculated numerically using~\eref{neel-ov}.

Figure~\ref{fig0:neel-ov} plots the squared overlaps $|\langle\lambda|N\rangle|^2$ 
between the eigenstates $|\lambda\rangle$ of the $XXX$ chain and the N\'eel state 
$|N\rangle$ for chains with $L=26,32,36,38$. Figure~\ref{fig0:neel-ov} (a) shows the 
squared overlaps versus the eigenstate energy density $E/L$. Clearly, the overlaps 
decay exponentially as a function of $L$, as expected. Moreover, for each $L$ 
we observe a rapid decay as a function of the eigenstate energy density. 
This decay is exponential, as shown in the inset of Figure~\ref{fig0:neel-ov} (a).

Complementary information is shown in Figure~\ref{fig0:neel-ov} (b) plotting the 
the histograms of $-2\log|\langle\lambda|N\rangle|/L$ (overlap distribution 
function). Notice that $2\log|\langle\lambda|N\rangle|/L$ is the driving 
term in the Quench Action approach. The factor $1/L$ in the definition takes 
into account that all the overlaps vanish exponentially in the thermodynamic 
limit, or at least the overlaps that are relevant for the quench action. 
Notice that larger values of $-2\log|\langle\lambda|N\rangle|/L$, correspond 
to faster decaying, i.e., asymptotically for large $L$, smaller overlaps.

Figure~\ref{fig0:neel-ov} (b) demonstrates that the vast majority of the eigenstates 
have small N\'eel overlap. Specifically, a peak at $\sim 0.5$ is visible in the 
Figure, and a rapid decay at smaller values of ... .

Interestingly, the data suggest the behavior
%
\begin{equation}
|\langle\lambda|N\rangle|^2\propto\exp(-\kappa L),
\end{equation}
%
with $0.18\lesssim\kappa\lesssim0.7$. The vertical dash-dotted line in 
Figure~\ref{fig0:neel-ov} is the overlap between the ground state of the 
$XXX$ chain and the N\'eel state. This is obtained using the ground state 
root distribution $\rho_1(\lambda)\propto1/\cosh(\pi\lambda)$  
and~\eref{neel-ov-th}. The dashed line denotes the N\'eel overlap 
$\sim 2/B(L,L/2)$ with the $S_z=0$ component of the ferromagnetic state.

The vertical dotted line in the Figure shows the expected result for 
$-2\log|\langle\lambda|N\rangle|/L$ in the thermodynamic limit as 
obtained from the quench action.

This is obtained by usigng~\eref{neel-ov-th} and the saddle point root distributions 
$\rho_n^*$ (cf.~\eref{rho1-sp}-\eref{rho3-sp} for the results up to $n=3$). Note that 
its position does not coincide with the peak of the overlap distribution function, as 
expected. This is reflecting the competition between the driving term and the entropy 
term in the quench action.

@@

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel}
\end{center}
\caption{ Overlap sum rules for the Neel state: The role of the zero-momentum 
 strings. (a) The overlap sum rule $\sum_{\lambda}|\langle\lambda|N\rangle|^2=1$, 
 with $|N\rangle$ the Neel state and $|\lambda\rangle$ the eigenstates  of 
 the $XXX$ spin chain. The $x$-axis shows $1/L$, with $L$ the chain length. 
 The circles are Bethe ansatz results for chains up to $L=36$. The results 
 are obtained via a full scanning of the chain Hilbert space. Only the 
 eigenstates with no zero-momentum strings are considered. The dash-dotted 
 line is a fit to $A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Inset: 
 The same data as in the main Figure plotted versus $1/L^{1/2}$. (b) 
 The same as in (a) for the sum rule $\sum_{\lambda}|\langle
 \lambda|N\rangle|^2e(\lambda)=1/2$, with $e(\lambda)$  the 
 eigenstates energy density. 
}
\label{fig1:neel-sr}
\end{figure}
%##################################################################

Here we investigate the effect of the Hilbert space truncation removed the 
zero-momentum strings on the N\'eel sum rules. 
We focus on the ``trivial'' sum rule, i.e., the normalization of the initial 
state 
%
\begin{equation}
\label{sr-trivial}
\langle N|N\rangle=\sum\limits_{\lambda}|\langle\lambda|N\rangle|^2=1, 
\end{equation}
%
and the local conserved charges sum rules 
%
\begin{equation}
\label{sr-charge}
Q_n^{(0)}=\langle N|Q_n|N\rangle=\sum\limits_{\lambda}|\langle\lambda|N\rangle|^2
q_{n}(\lambda)\quad\textrm{with}\quad n\in\mathbb{N}, 
\end{equation}
%
where $Q_n$ are the conserved charges of the $XXX$ chain (cf. 
subsection~\ref{sec:1.5}) and $q_n(\lambda)$ are the charges eigenvalues over 
the generic Bethe state $|\lambda\rangle$ (cf.~\eref{qngnk} and~\eref{gnk}).

In~\eref{sr-charge} $Q_n^{(0)}$ is the charge expectation value over the initial 
state, i.e., the N\'eel state. These can be calculated as done in 
Ref.~\cite{fagotti-2013}.

Note that due to the locality of $Q_n$, the translational invariance of the initial 
state, and the periodic boundary conditions, the charge density $Q_n^{(0)}/L$ does 
not depend on the chain size. For the N\'eel state, in the thermodynamic limit 
$Q_n^{(0)}$ is obtained using~\eref{q0-th} and the quench action root distributions 
$\pmb{\rho}^*$ (cf.~\eref{rho1-sp}-\eref{rho3-sp}).

The truncation of the chain Hilbert space removing the zero-momentum strings has 
dramatic effects on the sum rules~\eref{sr-trivial} and~\eref{sr-charge}.

This is illustrated in Figure~\ref{fig1:neel-sr} (a) focusing on the ``trivial'' 
sum rule~\eref{sr-trivial} and in Figure~\ref{fig1:neel-sr} (b) for~\eref{sr-charge} 
with $n=2$, i.e., the energy sum rule. 
For the N\'eel state, for any finite $L$ Eq.~\eref{sum} gives $Q^{(0)}_2/L=-1/2$.

The circles in Figure~\ref{fig1:neel-sr} are the numerical results for the sum rules 
obtained via a full scanning of the Hilbert space of the Heisenberg chain, excluding 
the zero momentum strings. The data are plotted against the inverse chain length $1/L$, 
for $L\le 38$. The expected results including all the eigenstates are shown as 
horizonthal dotted lines in the Figure.

Clearly, both the sum rules are violated, due to the exclusion of the to zero-momentum 
strings. Moreover, in both panesl the data suggest a vanishing behavior upon increasing 
the chain size. In particular, the dash-dotted lines in the Figures are fits to  
$A/L^{1/2}+B/L$, with $A,B$ fitting parameters, which perfectly describes the behavior 
of the data.

The vanishing behavior as $\propto 1/L^{1/2}$ is reflecting the vanishing 
of the fraction of non-zero momentum string eigenstates in the thermodynamic limit.  
Precisely, from~\eref{zneel1} and~\eref{ztilde} one obtains 
%
\begin{equation}
\label{beh}
\frac{\widetilde Z_{Neel}}{Z_{Neel}}\propto\frac{4}{\sqrt{\pi L}}. 
\end{equation}
% 
in the thermodynamic limit.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel state overlap}
\label{sec:2.1}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The string hypothesis: Reduced N\'eel overlap formulas}
\label{sec:2.2}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overlap with the Majumdar-Ghosh state}
\label{sec:2.3}

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
\label{mg}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
The overlap between a generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state can be obtained from the N\'eel state overlap in Eq.~\eref{neel-ov} 
as~\cite{pozsgay-2014} 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\pm\lambda_j\}_{j=1}^m\rangle=\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)
\langle N|\{\pm\lambda_j\}_{j=1}^m\rangle
\end{equation}
%
The mutliplicative factor in~\eref{mg-ov}, using the string hypothesis is 
rewritten as 
%
\begin{eqnarray}
\label{mg-k}
\fl\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)=\\\nonumber\qquad
2^m\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}
\lambda_{j;\ell}^{1+(-1)^j}(\lambda^2_{j;\ell}+j^2)\prod\limits_{k=0}^{\lfloor 
j/2\rfloor}\Big[\lambda_{j;\ell}^2+\Big(2k+\frac{1-(-1)^j}{2}\Big)^2\Big]^{-2}
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The N\'eel overlap in the thermodynamic limit}
\label{sec:2.4}

@@

The functional integral in~\eref{qa-d-ensemble} can be evaluated in the limit 
$L\to\infty$ using the saddle point approximation. One has to minimize the 
functional ${\mathcal F}(\pmb{\rho})$ defined as 
%
\begin{equation}
L{\mathcal F}(\pmb{\rho})\equiv 2\Re\log\langle\pmb{\rho}|\Psi_0
\rangle+S_{YY}(\pmb{\rho}(\lambda))  
\end{equation}
% 
with respect to $\pmb{\rho}$, i.e., solving $\delta{\mathcal F}(\pmb{\rho})/\delta
\pmb{\rho}|_{\pmb{\rho}=\pmb{\rho^*}}=0$, under the constraint that the thermodynamic BGT 
equations~\eref{bgt-th} hold. Finally, one obtains from~\eref{qa-d-ensemble} that in the 
thermodynamic limit 
%
\begin{equation}
\label{obs-th}
\langle{\mathcal O}\rangle=\langle\pmb{\rho^*}|{\mathcal O}|\pmb{\rho^*}\rangle. 
\end{equation}
%
Remarkably, for the quench with initial state the N\'eel state $|\Psi_0\rangle=|N\rangle$ the 
saddle point root distributions ${\rho_n^*(\lambda)}_{n=1}^\infty$ can be obtained 
analytically~\cite{brockmann-2014}. The first few are given as 
%
\begin{eqnarray}
\label{rho1-sp}
\fl\quad\rho^*_1(\lambda)=\frac{8(4+\lambda^2)}{\pi(19+3\lambda^2)(1+6\lambda^2+
\lambda^4)}\\\label{rho2-sp}
\fl\quad\rho_2^*(\lambda)=\frac{8\lambda^2(9+\lambda^2)(4+3\lambda^2)}{\pi(2+\lambda^2)
(16+14\lambda^2+\lambda^4)(256+132\lambda^2+9\lambda^4)}\\\label{rho3-sp}
\fl\quad\rho_3^*(\lambda)=\frac{8(1+\lambda^2)^2(5+\lambda^2)(16+\lambda^2)(21+\lambda^2)}
{\pi(19+3\lambda^2)(9+624\lambda^2+262\lambda^4+32\lambda^6+\lambda^8)
(509+5\lambda^2(26+\lambda^2))}.
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{N\'eel overlaps: The role of the zero-momentum strings Bethe states}
\label{sec:5}

In this section we discuss generic features of the overlaps between the 
eigenstates (Bethe states) of the Heisenberg spin chain and the N\'eel state 
(cf.~\eref{neel}). We exploit the Bethe ansatz solution of the chain (see 
section~\ref{sec:1}) as well as exact results for the N\'eel overlaps (see 
section~\ref{sec:2}). We focus on finite chains with $L\lesssim 40$ sites, 
considering, for any $L$, the full Hilbert space of the chain. 
The Bethe states having, in principle, non zero N\'eel overlap are the 
so-called parity-invariant Bethe states (see section~\ref{sec:2}). We 
denote their total number as $Z_{Neel}$. Crucially, here we restrict 
ourselves to the subset of parity-invariant Bethe states corresponding 
to solutions of the Bethe-Gaudin-Takahashi (BGT) equations~\eref{bgt-eq} 
that do not contain zero-momentum strings. We denote the total number of 
these eigenstates as $\widetilde Z_{Neel}$. Both $Z_{Neel}$ and $\widetilde 
Z_{Neel}$ are given in terms of the chain length $L$ by simple combinatorial 
formulas that we provide.

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel_overlaps}
\end{center}
\caption{ N\'eel overlaps with the eigenstates of the Heisenberg spin 
 chain: Numerical results obtained from the full scanning of the chain 
 Hilbert space. Eigenstates corresponding to zero-momentum strings are 
 escluded. (a) Squared overlaps $|\langle\lambda|N\rangle|^2$ plotted as 
 function of the eigenstates energy density $E/L$. Here $|\lambda\rangle$ 
 denotes the generic eigenstate. The data are for chains with length $26
 \le L\le 38$. The inset is to highlight the exponential decay as a 
 function of $E/L$. Note the logarithmic scale on the $y$ axis. (b) 
 Overlap distribution function: Histograms of $-2\log\langle\lambda|N\rangle/L$. 
 The $y$-axis is rescaled by a factor $10^5$ for convenience. The dash-dotted 
 and dashed vertical lines are the N\'eel overlaps with the $XXX$ chain ground 
 state and the ferromagnetic state, respectively. The dotted line (see 
 the arrow) is the result obtained using the quench action approach. 
}
\label{fig0:neel-ov}
\end{figure}
%##################################################################

Interestingly, the fraction of eigenstates with no zero-momentum strings, i.e.,  
$\widetilde Z_{Neel}/Z_{Neel}$, is vanishing as $L^{-1/2}$ in the thermodynamic 
limit, meaning that zero-momentum strings eigenstates are dominant, at least 
in number, for large chains. This has dramatic consequences, for instance, at 
the level of the overlap sum rules for the local conservation laws of the model. 
Precisely, although the sum rules are fixed by the initial (N\'eel) state 
expectation value of the local charges, we observe striking violations for 
any finite chain. Moreover, all the sum rules exhibit vanishing behavior as 
$L^{-1/2}$ upon increasing the chain size, reflecting the same vanishing behavior 
as $\widetilde Z_{Neel}/Z_{Neel}$. A similar scenario holds for the overlaps with 
the Majumdar-Ghosh state, where, however, excluding the zero-momentum strings leads 
to a $1/L$ behavior. This demonstrates that for finite chains the physical effects 
of zero-momentum strings eigenstates are not negligible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel overlap distribution function: Overview}
\label{sec:5.1}

Here we overview the Bethe ansatz results for the N\'eel overlaps with the 
eigenstates of the $XXX$ chain. The total number of partiy-invariant 
eigenstates $Z_{Neel}$ having, in principle, non-zero N\'eel overlap is given 
as 
%
\begin{equation}
\label{zneel1}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1, 
\end{equation}
%
with $B(n,m)\equiv n!/(m!(n-m)!)$ the Newton binomial. The proof of~\eref{zneel1} 
is obtained by counting all the parity-invariant BGT quantum number configurations, 
and it is reported in~\ref{app-1}. Note that $Z_{Neel}$ provides only an upper 
bound for the number of eigenstates with non-zero N\'eel overlap, as it is clear 
from the exact diagonalization results shown in Table~\ref{table:neel}. This is 
because parity-invariant eigenstates with a single zero-momentum even-length 
string, which are included in~\eref{zneel1}, have identically zero N\'eel 
overlap~\cite{brockmann-2014}. This is not related to the symmetries of the N\'eel 
state, but to the ``accidental'' vanishing of the prefactor in the overlap 
formula~\eref{neel-ov}. Finally, after excluding the zero-momentum strings 
eigenstates, the total number of remaining eigenstates $\widetilde Z_{Neel}$, 
which are the ones considered here, is given as (see~\ref{app-2} for the proof) 
%
\begin{equation}
\label{ztilde}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big).
\end{equation}
%
An overview of generic features of the overlaps is given in Figure~\ref{fig0:neel-ov} 
(a) plotting the squared N\'eel overlaps $|\langle\lambda|N\rangle|^2$ with the $XXX$ 
chain eigenstates $|\lambda\rangle$ versus the energy density $E/L$. The figure shows 
results for chains with $26\le L\le 38$ sites. The data are obtained by generating all 
the relevant parity-invariant BGT quantum numbers, and solving the associated BGT 
equations~\eref{bgt-eq}, to obtain the rapidities of $XXX$ chain eigenstates. The 
N\'eel overlaps are calculated numerically using~\eref{neel-ov}. Note that for $L=38$ 
from~\eref{ztilde} the total number of overlap shown in the Figure is $\widetilde Z_{Neel}
\sim 10^5$.

Clearly, the overlaps decay exponentially as a function of $L$, as expected. Moreover, 
at each finite $L$ a rapid decay as a function of $E/L$ is observed. The inset of 
Figure~\ref{fig0:neel-ov} (a) suggests that the decay is exponential. Complementary 
information about the overlaps is shown in Figure~\ref{fig0:neel-ov} (b) plotting the 
histograms of $\kappa\equiv-2\log|\langle\lambda|N\rangle|/L$ (overlap distribution 
function). Larger values of $\kappa$, correspond to a faster decay with $L$ of the 
overlaps. The factor $1/L$ in the definition takes into account that the N\'eel 
overlaps vanish exponentially as $|\langle\lambda|N\rangle|^2\propto e^{-\kappa L}$ in 
the thermodynamic limit. Note that $\kappa$ is the driving term in the quench action 
approach (cf~\eref{neel-ov-th}). As expected, from Figure~\ref{fig0:neel-ov} (b) one 
has that the majority of the $XXX$ chain eigenstates give small N\'eel overlap (note 
the maximum at $\kappa\sim 0.5$). Interestingly, the data suggest that with 
$0.18\lesssim\kappa\lesssim0.7$. The vertical dash-dotted line in the figure is 
the $\kappa$ obtained from the N\'eel overlap of the ground state of the $XXX$ 
chain in the thermodynamic limit. This is derived using the ground state root 
distribution $\rho_1(\lambda)\propto1/\cosh(\pi\lambda)$ and~\eref{neel-ov-th}. 
On the other hand, the vertical dashed line denotes the N\'eel overlap $\sim 2/B(L,L/2)$ 
of the $S_z=0$ component of the ferromagnetic multiplet, which is at the top of the
chain energy spectrum. Finally, the vertical dotted line in Figure~\ref{fig0:neel-ov} 
(b) shows the quench action result for $\kappa$ in the thermodynamic limit. This is 
obtained by using~\eref{neel-ov-th} and the saddle point root distributions $\rho_n^*$ 
(cf.~\eref{rho1-sp}-\eref{rho3-sp} for the results up to $n=3$). Note that $\kappa$ 
does not coincide with the peak of the overlap distribution function, as expected, 
due to the effect of the driving term~\eref{neel-ov-th} in the quench action approach.

@@

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel}
\end{center}
\caption{Overlap sum rules for the N\'eel state $|N\rangle$: The role of 
 the zero-momentum strings. (a) The overlap sum rule $\sum_{\lambda}|
 \langle\lambda|N\rangle|^2=1$. Here $|\lambda\rangle$ are the eigenstates  
 of the $XXX$ chain. The $x$-axis shows the inverse chain length $1/L$. 
 The circles are Bethe ansatz results for chains up to $L=38$. The data  
 are obtained by  a full scanning of the chain Hilbert space. Eigenstates  
 corresponding to zero-momentum strings are excluded. The dotted line is 
 the expected result at any $L$. The data are compatible with a vanishing 
 behavior in the thermodynamic limit. The dash-dotted line is a fit to 
 $A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Inset: The same data as 
 in the main Figure now plotted versus $1/L^{1/2}$. (b) 
 The same as in (a) for the energy sum rule $\sum_{\lambda}|\langle
 \lambda|N\rangle|^2Q_2(\lambda)=Q_2^{(0)}$, with $Q_2(\lambda)$ the 
 energy of the eigenstate $|\lambda\rangle$ and $Q_2^{(0)}/L=-1/2$ the 
 N\'eel state energy density (dotted line in the Figure). 
}
\label{fig1:neel-sr}
\end{figure}
%##################################################################

Here we illustrate the effect of the zero-momentum strings eigenstates on the 
N\'eel overlap sum rules. We focus on the ``trivial'' sum rule, i.e., the 
normalization of the N\'eel state  
%
\begin{equation}
\label{sr-trivial}
\langle N|N\rangle=\sum\limits_{\lambda}|\langle\lambda|N\rangle|^2=1. 
\end{equation}
%
We also consider the N\'eel expectation value of the local conserved charge 
$Q_n$ of the $XXX$ chain (see subsection~\ref{sec:1.5}), which provide 
the additional sum rules
%
\begin{equation}
\label{sr-charge}
Q_n^{(0)}=\langle N|Q_n|N\rangle=\sum\limits_{\lambda}|\langle\lambda|N\rangle|^2
Q_{n}(\lambda)\quad\textrm{with}\quad n\in\mathbb{N}, 
\end{equation}
%
where $Q_n(\lambda)$ are the charges eigenvalues over the generic Bethe state 
$|\lambda\rangle$ (cf.~\eref{qngnk} and~\eref{gnk}). 
In both~\eref{sr-trivial} and~\eref{sr-charge} the sum are restricted to the 
eigenstates with containing no zero-momentum strings. 
In~\eref{sr-charge} $Q_n^{(0)}$ is the charge expectation value over the initial 
state, i.e., the N\'eel state, which has been calculated in Ref.~\cite{fagotti-2013}. 
Due to the locality of $Q_n$, the translational invariance of the initial 
state, and the periodic boundary conditions, the density $Q_n^{(0)}/L$ does 
not depend on the chain size. For the N\'eel state, $Q_n^{(0)}$ can be calculated 
directly in the thermodynamic limit using~\eref{q0-th} and the root distributions 
$\pmb{\rho}^*$ (cf.~\eref{rho1-sp}-\eref{rho3-sp}).

The sum rules~\eref{sr-trivial} and~\eref{sr-charge} (for $n=2$, i.e., the energy 
sum rule), are shown in Figure~\ref{fig1:neel-sr} (a) and (b), respectively. 
Note that $Q^{(0)}_2/L=-1/2$ in~\eref{sr-charge} (horizonthal dotted line in 
Figure~\ref{fig1:neel-sr} (b)). 
The circles in Figure~\ref{fig1:neel-sr} (a) are the Bethe ansatz results excluding 
the zero momentum strings. The data are the same as in Figure~\ref{fig0:neel-ov}. 
The sum rules are plotted against the inverse chain length $1/L$, for $L\le 38$.

Clearly, both the sum rules are violated, due to the exclusion of the to zero-momentum 
strings. Moreover, in both Figure~\ref{fig1:neel-sr} (a) and (b) the data suggest a 
vanishing behavior upon increasing $L$. The dash-dotted lines in the Figures are fits to  
$A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Interestingly, the behavior as $\propto 
L^{-1/2}$ of the sum rules reflects that of the fraction of non-zero momentum string 
eigenstates in the thermodynamic limit $\widetilde Z_{Neel}/Z_{Neel}$. Specifically, 
from~\eref{zneel1} and~\eref{ztilde} it is straightforward to derive that in the 
thermodynamic limit 
%
\begin{equation}
\label{beh}
\frac{\widetilde Z_{Neel}}{Z_{Neel}}\propto\frac{4}{\sqrt{\pi L}}. 
\end{equation}
%
